DNNNNR0: How to Avoid OOTA Without Really Trying

### 2025-03-27 paper/rebuttal-mods.txt

x	Add a forward reference to Section 4.2 ("Global Optimization
	Can Destroy Dependencies") and the fact that this defeats the
	approach of ordering relaxed loads before relaxed stores.

x	Be less condescending towards Mark Batty.  Maybe
	s/is persisting with/is working on/.

x	Add a Section 1.3 ("Our Approach"?) that summarizes the
	rfe/sdep/non-sdep-dep approach and emphasizes proving
	OOTA cannot happen as opposed to an exact delineation
	between sdep and non-sdep.

x	Add citations to later sections that discuss litmus tests
	that are similar to those in the literature.

o	Consider adding a 2.2.x section that discusses the difficulties
	inherent in computing sdep for function calls, especially when
	the function will be part of a dynamic library that has not yet
	been written.  (This addresses Reviewer #72F's call for handling
	examples in Section 5 of [5].)  Or put this in a later section
	demonstrating how we handle the various litmus tests.

x	In section 2.2 where we start to talk about semantic dependencies
	in detail, we could mention syntactic dependencies too.  Explain
	what they are and how they differ from semantic dependencies.
	Then mention that some work in the literature (such as parts of
	[5]) is not always careful to distinguish clearly between the
	two forms.  Maybe have an entire subsection on this topic.

@@

x	In Section 4.2 ("Global Optimization Can Destroy Dependencies"),
	note that this optimization breaks the approach of ordering
	relaxed loads before relaxed stores.

x	In section 6.1, define "witness" informally as "another execution
	in which the load(s) obtain a different value but all else is
	the same as far as possible, and yet the store acts differently".

o	Add diagrams illustrating what the various components of sections
	6.1 and 6.3 refer to. (Reviewer #72B)

	Brainstorming for diagrams:

	o	Section 1.3: Showing that an OOTA cycle cannot close.
		Time increases downwards.

	o	Section 6.1: Diagram showing source-code plane
		and object-code plane, with mapping of accesses and
		dependencies from one to the other.  Semantic source-level
		dependencies are always mapped, but non-semantic
		dependencies might or might not be mapped.

		Section 6.1: Show the witness.	Or should this instead
		be a side-by-side listing?

	o	Section 6.3: Also show the witness.  Maybe only one
		witness diagram/listing?

@@

o	Review Sections 6.1 and 6.3 for clarity per Reviewer #72F.

x	On page 9, "While it possible".  [ Since fixed by Alan. ]

x	In Section 7.4 ("Merging Quasi-Volatile Loads"), change focus
	from intuitive notions to a community-process approach for
	determining what optimizations are legitimate.  Math cannot
	handle this.  We avoid the problem by looking at the compiler
	implementation, which is downstream of the community process.

x	Also in Section 7.4 change "our intuitive notions" to "one's
	intuitive notions".

@@

o	Add a section before Section 7 ("Issues and Refinements") that
	shows how our approach handles a few well-known examples of OOTA.
	This needs to be in greater detail than in the "Causality Test
	Cases" discussions in P3064R2 ("How to Avoid OOTA Without Really
	Trying").

x	Possible future work: Relax quasi-volatile semantics to permit
	load speculation, for example, use of the cmov instruction in
	cases where the source code conditionally executes the atomic
	load.

### 2025-04-30 paper/ootavacuous.tex
ootavacuous: Add note from search for ++ litmus tests
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>
ootavacuous: Add a section on semantic dependencies and function calls
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

@@

Consider the following example, with variable \co{i} initially zero:
% @@@ Citation?
\begin{quote}
\begin{verbatim}
int foo(int a, int b)
{
   return a / b;
}

>>>>

\subsubsection{Semantic Dependencies and Function Calls}
\label{sec:Semantic Dependencies and Function Calls}

Boehm and Dempsky~\cite[Section 5]{Boehm:2014:OGA:2618128.2618134}
point out that arbitrary function calls can pose challenges for any
attempt to determine semantic dependencies based solely on analysis
of the source code, using the following example:
\begin{quote}
\begin{verbatim}
y = f(x);
\end{verbatim}
\end{quote}
And it is indeed impossible to determine whether or not there is a
semantic dependency from \co{x} to \co{y} without reference to the
definition of \co{f()}.
For some definitions of \co{f()}, a semantic dependency exists, and for
others it does not.

Note that \co{f} might well be a pointer to a function, in which case
the existence of that semantic dependency might well vary at runtime as
the value of function pointer \co{f} changes.

@@

Consider the following example, with variable \co{i} initially zero:
% @@@ Citation?
% @@@ "Outlawing ghosts: Avoiding out- of-thin-air results" has a
% @@@    ++ example, but is concerned about hoisting that before
% @@@    an "if" statement.
\begin{quote}
\begin{verbatim}
int foo(int a, int b)
{
   return a / b;
}
